---
title: "Urban refugia enhance persistence of an endemic keystone species facing a rapidly spreading invasive predator"
author: "Marc Vez-Garzón, Sandra Moreno, Guillem Casbas, Victor Colomar, Oriol Lapiedra"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
editor: visual
---

# Introduction

This R script collects all the steps followed to achieve the results shown in the article.

# Load required packages

```{r packages}
#| warning: false
library(raster)
library(readr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(EnvStats)
library(sf)
library(ggpubr)
library(extRemes)
library(lme4)
library(performance)
library(glmmTMB)
library(DHARMa)
library(pgirmess)
```

# Ibiza map

## Load map and define sampling points

```{r Ibiza map sampling points}
#| warning: false

# Map working directory 
setwd("C:/Users/marc9/Desktop/Marc/CREAF/Urban refugia")

# Load 10x10m raster map
Eivissa_map <- raster("Eivissa2021.tif")

# Load data
Urban_Refugia_MVG_cens_total_net <- read_csv("UR_new_classifcation_nov_2023_ibiza_inv.csv")

# Sampling points coordinates (long/lat)
coords <- Urban_Refugia_MVG_cens_total_net %>%
  dplyr::select(Latitude, Longitude)

# Transform to correct format 
coordinates(coords) <- ~ Longitude + Latitude

# 50 metres buffer around each sampling point
buff_shp <- buffer(coords, 50)

# Plot Ibiza map (no roads, but with sampling points)
colors2 <- c("#89d2a3", "#a7dca5", "#c5ecac", "#fbdf9d", "#e58087", "#d1d1d1", "#dceef3", "#8dc8d8", "#d4bbfc")

buff_shp_plot <- buffer(coords, 200)

plot(Eivissa_map, axes = T, col = colors2, legend = F, xaxt = "n", yaxt = "n")

plot(buff_shp_plot, add = TRUE)
```

```{r plot Ibiza roads}
# Plot Ibiza map (with roads)
# Roads file
roads <- read_sf(dsn = "C:/Users/marc9/Desktop/Marc/CREAF/Invasion snake/Final_project", layer = "carreteras_ibiza")

roads$clased <- factor(roads$clased, levels = c("Carretera convencional", "Carretera multicarril", "Urbano"))

roads_subset <- subset(roads, clased %in% c("Carretera convencional", "Carretera multicarril", "Urbano"))

colores <- c(rep("black", 3))

tamanios <- c(1, 1, 0.05)

plot(Eivissa_map, axes = T, col = colors2, legend = F, xaxt = "n", yaxt = "n")

for (i in 1:length(levels(roads_subset$clased))) {
  subset_roads <- roads_subset[roads_subset$clased == levels(roads_subset$clased)[i], ] 
  lines(subset_roads, lwd = tamanios[i], col = colores[i]) 
}
```

## Calculate urbanization index

```{r urbanization index}
#| warning: false

# Extract use of land information from every 10x10m cell inside the 50m buffer
landcover_eivissa <- raster::extract(Eivissa_map, coords, buffer = 50, df=TRUE)

# Convert to data frame (ID = sampling point | Map = land use category)
landcover_eivissa <- as.data.frame(landcover_eivissa)

# Summarise by sampling point and land use category (how many 10x10m cells of each type are within each buffer area)
landcover_eivissa <- landcover_eivissa %>% 
  group_by(ID, Map) %>%
  dplyr::summarise(n=n())

# Total number of 10x10m cells within each buffer
landcover_eivissa <- landcover_eivissa %>% 
  group_by(ID) %>%
  dplyr::mutate(total=sum(n))

# Percentage of cells from each category with respect to the total number of cells within each buffer
landcover_eivissa <- landcover_eivissa %>% 
  group_by(ID, Map) %>%
  dplyr::summarise(value=n/total)

# Wide format
landcover_eivissa <- spread(landcover_eivissa, Map, value)

# Name of each land use category
landcover_eivissa <- landcover_eivissa %>%
  dplyr::rename(Tree_cover="10",
                Shrubland="20",
                Grassland="30",
                Cropland="40",
                Built_up="50",
                Bare.sparse_vegetation="60",
                Permanent_water_bodies="80")

# Add Place, Site and Project columns
landcover_eivissa <- landcover_eivissa %>%
  add_column(Place = Urban_Refugia_MVG_cens_total_net$Place,
             Site = Urban_Refugia_MVG_cens_total_net$Site_ID,
             Project = Urban_Refugia_MVG_cens_total_net$Project)

# Place, Site and Project columns to the beginning of the data frame
landcover_eivissa <- landcover_eivissa %>%
  relocate(c(Place, Site, Project), .after = ID)

# NA as 0
landcover_eivissa[is.na(landcover_eivissa)] <- 0

# Urbanization index (Build/Total)
landcover_eivissa$Urbanization_index <- landcover_eivissa$Built_up/
  (landcover_eivissa$Tree_cover + 
     landcover_eivissa$Shrubland + 
     landcover_eivissa$Grassland + 
     landcover_eivissa$Cropland + 
     landcover_eivissa$Bare.sparse_vegetation + 
     landcover_eivissa$Built_up)

```

# Lizard censuses

## Data set cleaning and exploratory plots

```{r data set filtering}
#| warning: false

# Select desired columns
Urban_Refugia_MVG_seleccio <- Urban_Refugia_MVG_cens_total_net[, c(1, 3, 4, 6, 41, 11:26, 37, 39, 40, 42)]

# Convert factor variables to factor
Urban_Refugia_MVG_seleccio$Place <- as.factor(Urban_Refugia_MVG_seleccio$Place)
Urban_Refugia_MVG_seleccio$Project <- as.factor(Urban_Refugia_MVG_seleccio$Project)
Urban_Refugia_MVG_seleccio$Date <- as.factor(Urban_Refugia_MVG_seleccio$Date)
Urban_Refugia_MVG_seleccio$Site_ID <- as.factor(Urban_Refugia_MVG_seleccio$Site_ID)
Urban_Refugia_MVG_seleccio$Inv_state <- as.factor(Urban_Refugia_MVG_seleccio$Inv_state)
Urban_Refugia_MVG_seleccio$Classifica <- as.factor(Urban_Refugia_MVG_seleccio$Classifica)


# Total number of Podarcis per replicate
Urban_Refugia_MVG_seleccio$Suma_total <- Urban_Refugia_MVG_seleccio$`Total Male` +  
  Urban_Refugia_MVG_seleccio$`Total Fema` + 
  Urban_Refugia_MVG_seleccio$`Total Juve` + 
  Urban_Refugia_MVG_seleccio$`Total NA`

# Reorder variables
Urban_Refugia_MVG_seleccio <- Urban_Refugia_MVG_seleccio[, c(1:21, 26, 22:25)]

# Merge urbanization index to data frame
Urban_Refugia_MVG_seleccio$Urbanization_index <- landcover_eivissa$Urbanization_index

# Order of factors
Urban_Refugia_MVG_seleccio$Inv_state <- factor(Urban_Refugia_MVG_seleccio$Inv_state,
                                               levels = c("NonInvaded", "Invaded"))

# New variable that merges invasion state and habitat type
Urban_Refugia_MVG_seleccio$Code <- paste0(Urban_Refugia_MVG_seleccio$Inv_state, "_", Urban_Refugia_MVG_seleccio$Project)

# Set theme
theme_set(theme_classic() + theme(legend.position = 'bottom') + theme(text = element_text(size = 17)))

# Lizard abundance plot
ggplot(Urban_Refugia_MVG_seleccio, aes(x = factor(Code, level=c('NonInvaded_PeriUrban', 'NonInvaded_Urban', 'Invaded_PeriUrban', "Invaded_Urban")), y = Suma_total, fill = Code)) +
  geom_violin(trim = TRUE, adjust = 1.5, width = 0.75) +  
  labs(x = "\nInvasion and urbanization status", y = "Lizard abundance") +
  theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5), legend.position = "none") +
  geom_jitter(shape = 16, position = position_jitterdodge(jitter.width = 0.6, dodge.width = 0.75, jitter.height = 0.2)) +  
  stat_summary(fun.y = mean, geom = "point", shape = 21, col = "black", fill = "white", stroke = 1, size = 4, position = position_dodge(width = 0.7)) +  
  scale_fill_manual(values = c( "grey60", "grey30", "tomato1", "red3"),
                    name = "Treatment", 
                    breaks=c('NonInvaded_PeriUrban', 'NonInvaded_Urban', 'Invaded_PeriUrban', "Invaded_Urban")) +
  scale_x_discrete(labels=c("Non-invaded\nPeri-urban", "Non-invaded\nUrban", "Invaded\nPeri-urban", "Invaded\nUrban")) +
  stat_n_text(y.pos = NULL, color = "black", text.box = FALSE)

# Mean and sd for each treatment
aggregate(Suma_total ~ Inv_state + Project, Urban_Refugia_MVG_seleccio, function(x) c(mean = mean(x), sd = sd(x)))

# Select urban sampling points
Urban_Refugia_MVG_seleccio_urban <- Urban_Refugia_MVG_seleccio[Urban_Refugia_MVG_seleccio$Project == "Urban", ]

# Urbanization index distribution (urban)
urb_urb_complete <- ggplot(Urban_Refugia_MVG_seleccio_urban, aes(x = `Index Urb`)) + 
  geom_density(fill = "grey20", alpha = 0.6, linewidth = 1) +
  ylab("Density") +
  xlab("Urbanization index \n(Urban sites)") +
  theme(
    axis.title.y = element_text(vjust = 2),
    axis.title.x = element_text(vjust = -0.5)) 

# Select peri-urban sampling points
Urban_Refugia_MVG_seleccio_periurban <- Urban_Refugia_MVG_seleccio[Urban_Refugia_MVG_seleccio$Project == "PeriUrban", ]

# Urbanization index distribution (peri-urban)
urb_periurb_complete <- ggplot(Urban_Refugia_MVG_seleccio_periurban, aes(x = `Index Urb`)) + 
  geom_density(fill = "grey80", alpha = 0.6, linewidth = 1) +
  xlim(0, 1) + 
  ylab("Density") +
  xlab("Urbanization index \n(Peri-urban sites)") +
  theme(
    axis.title.y = element_text(vjust = 2),
    axis.title.x = element_text(vjust = -0.5))

# Plot urbanization index
ggarrange(urb_periurb_complete, urb_urb_complete,  ncol = 2, nrow = 1)
```

## One-way ANOVA + Tukey's HSD test on lizard abundance

```{r ANOVA + Tukey lizard abundance}
#| warning: false

# One-way ANOVA test
summary(aov(Suma_total ~ Code, data = Urban_Refugia_MVG_seleccio))

# Tukey's HSD test
TukeyHSD(aov(Suma_total ~ Code, data = Urban_Refugia_MVG_seleccio))

```

# Zero-inflated Poisson GLMM

## Data preparation

```{r ZIP preparation}
#| warning: false

# Percentage of 0 in our data
length(Urban_Refugia_MVG_seleccio$Suma_total[Urban_Refugia_MVG_seleccio$Suma_total == 0])/
  nrow(Urban_Refugia_MVG_seleccio) * 100

# Year of invasion index (0: not-invaded / 1: oldest invaded site)
Urban_Refugia_MVG_seleccio$`3P 250 200` <- Urban_Refugia_MVG_seleccio$`3P 250 200` - max(Urban_Refugia_MVG_seleccio$`3P 250 200`)

diff_years <- max(Urban_Refugia_MVG_seleccio$`3P 250 200`) - min(Urban_Refugia_MVG_seleccio$`3P 250 200`)

Urban_Refugia_MVG_seleccio$`3P 250 200` <- abs(Urban_Refugia_MVG_seleccio$`3P 250 200`)

Urban_Refugia_MVG_seleccio$`3P 250 200` <- Urban_Refugia_MVG_seleccio$`3P 250 200`/diff_years

# Initial GLMM (Poisson)
P5 <- glmer(Suma_total ~ `3P 250 200` + Urbanization_index + (1 | Place/Site_ID), 
            family = poisson(), 
            data = Urban_Refugia_MVG_seleccio, 
            verbose = FALSE)

summary(P5)

# Overdispersion check: No overdispersion (no need to use negative binomial)
check_overdispersion(P5)

# Zero inflation check: zero-inflation detected
check_zeroinflation(P5, tolerance = 0.05)
```

## Zero-inflated Poisson

```{r ZIP}
#| warning: false

# Zero-inflated Poisson
ZIP3_Year_Urb <- glmmTMB(
  Suma_total ~ `3P 250 200` + Urbanization_index + (1 | Place/Site_ID),
  ziformula = ~ `3P 250 200` + Urbanization_index, 
  family = poisson(link = "log"), 
  data = Urban_Refugia_MVG_seleccio
)

summary(ZIP3_Year_Urb)

# Zero-inflated Poisson results (table)
sjPlot::tab_model(ZIP3_Year_Urb,
                  show.aicc = TRUE, 
                  show.obs = FALSE, 
                  show.ngroups = FALSE, 
                  p.style = "stars",
                  transform = NULL,
                  show.intercept = TRUE,
                  dv.labels = c("ZIP"),
                  string.p = "p-value",
                  string.est = "Incidence Rate Ratios")

# Data frame preparation for results visualization
data_ZIP3_Year_Urb <- data.frame(
  Model  = as.factor(c("Poisson", "Poisson", "Poisson", "Zero_inf", "Zero_inf", "Zero_inf")),
  Variable = as.factor(c("Urb_index", "Year_inv", "Intercept", "Urb_index", "Year_inv", "Intercept")),
  Value = c(0.75, -2.68, 0.33, -1.82, 4.53, -0.91),
  Lower = c(0.06, -4.32, -0.41, -3.50, 2.29, -1.63),
  Upper = c(1.44, -0.64, 1.06, -0.14, 6.76, -0.19),
  pvalue = as.factor(c("< 0.05", "< 0.05", "> 0.05", "< 0.05", "< 0.001", "< 0.05")))

dotCOLS = c("grey80","tomato1")
barCOLS = c("grey30","red3")

# Zero-inflated Poisson model plot
ggplot(data_ZIP3_Year_Urb, aes(x=Variable, y=Value, ymin=Lower, ymax=Upper, col=Model, fill=Model)) + 
  geom_linerange(size=5,position=position_dodge(width = 0.5)) +
  geom_hline(yintercept=0, lty=2) +
  geom_point(size=3, shape=21, colour="black", stroke = 0.5, position=position_dodge(width = 0.5)) +
  scale_fill_manual(values=barCOLS, labels = c("Poisson", "Zero-inflated")) +
  scale_color_manual(values=dotCOLS, labels = c("Poisson", "Zero-inflated")) +
  scale_x_discrete(name="Fixed factors", limits=rev,  labels=c('Year of\ninvasion', 'Urbanization\n index', 'Intercept')) +
  scale_y_continuous(name="Odds ratio", limits = c(min(data_ZIP3_Year_Urb$Lower), max(data_ZIP3_Year_Urb$Upper))) +
  expand_limits(y=c(0.1, 50)) + 
  guides(color = guide_legend(reverse=TRUE), fill = guide_legend(reverse=TRUE)) +
  coord_flip() +
  theme(legend.position="bottom") +
  theme(text = element_text(size = 17),
        axis.title.y = element_text(vjust = 2))
```

## Zero-inflated + Poisson part ANOVAs

```{r ZIP ANOVA}
#| warning: false

# Zero-inflated part's ANOVA
glmmTMB:::Anova.glmmTMB(ZIP3_Year_Urb, 
                        type = 2, 
                        test.statistic = "Chisq", 
                        component = "zi")

# Poisson part's ANOVA
glmmTMB:::Anova.glmmTMB(ZIP3_Year_Urb, 
                        type = 2, 
                        test.statistic = "Chisq", 
                        component = "cond")
```

# Spatial autocorrelation

```{r spatial autocorrelation}
#| warning: false

# Extract residuals
res_ZIP3 <- as.data.frame(residuals(ZIP3_Year_Urb))

colnames(res_ZIP3) <- c("residuals_P3")

# Add columns of interest
res_ZIP3 <- res_ZIP3 %>%
  add_column(Place = Urban_Refugia_MVG_cens_total_net$Place,
             Site = Urban_Refugia_MVG_cens_total_net$Site_ID,
             Project = Urban_Refugia_MVG_cens_total_net$Project,
             Latitude = Urban_Refugia_MVG_seleccio$Latitude,
             Longitude = Urban_Refugia_MVG_seleccio$Longitude)

# Move the new columns to the beginning
res_ZIP3 <- res_ZIP3 %>%
  relocate(c(Place, Site, Project, Latitude, Longitude), .before = residuals_P3)

# Delete duplications
res_ZIP3 <- res_ZIP3 %>% group_by(Place, Site, Latitude, Longitude) %>%
  dplyr::summarise(across(c(residuals_P3), list(mean)))

# Spatial autocorrelation visualization
correlog_ZIP3 <- pgirmess::correlog(res_ZIP3[, c(4, 3)], res_ZIP3$residuals_P3_1, method="Moran", nbclass = NULL)

plot(correlog_ZIP3)


# Distance matrix between sampling points
dist_matrix <- dist(res_ZIP3[, c(3, 4)], diag = T, upper = T)

dist_matrix <- as.data.frame(as.matrix(dist_matrix))

# Moran's I test for distance-based autocorrelation
testSpatialAutocorrelation(res_ZIP3$residuals_P3_1,
                           distMat = dist_matrix)
```

# Urban filtering

## Load data and exploratory plots

```{r urban filter data + plots}
#| warning: false

# Load snake captures
Traps_snake <- read_delim("Traps_snake.csv", 
                          delim = ";", escape_double = FALSE, trim_ws = TRUE)

# Manual data set, in case file is not available
Zones <- c(rep("Outer", 18), rep("Intermediate", 8), rep("Inner", 15))

Total <- c(11, 2, 6, 4, 9, 4, 4, 7, 3, 2, 2, 1, 5, 2, 0, 0, 0, 0, 2, 0, 0, 4, 1, 1, 3, 2, rep(0, 15))

# Create data frame
Traps_snake <- data.frame(Zones = Zones, Total = Total)

# Mean and sd per treatment
media_por_zona <- aggregate(Total ~ Zones, data = Traps_snake, FUN = mean)
total_por_zona <- aggregate(Total ~ Zones, data = Traps_snake, FUN = length)

# Set theme
colors <- c("Outer" = "grey80", "Intermediate" = "grey50", "Inner" = "grey20")

theme_set(theme_classic() + theme(legend.position = 'bottom') + theme(text = element_text(size = 15)))

# Convert column "Zones" into factor
Traps_snake$Zones <- factor(Traps_snake$Zones, levels = c("Outer", "Intermediate", "Inner"))

# Snake captures plot
ggplot(Traps_snake, aes(x = Zones, y = Total, fill = Zones)) +
  geom_violin(trim = TRUE, adjust = 1.5, width = 0.75) +
  geom_jitter(shape = 16, size = 2, position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.75, jitter.height = 0.2)) +
  geom_point(data = media_por_zona, aes(y = Total), color = "black", fill = "white", size = 4, shape = 21) +
  labs(x = "",
       y = "Captured snakes per trap") +
  scale_fill_manual(values = colors) +
  scale_x_discrete(labels=c("Outer area", "Intermediate area", "Inner area")) +
  stat_n_text(y.pos = NULL, color = "black", text.box = FALSE) +
  theme(legend.position = "none",
        axis.ticks.x = element_blank())

aggregate(. ~ Zones, Traps_snake, function(x) c(mean = mean(x), sd = sd(x)))

```

## One-way ANOVA + Tukey's HSD test on snake captures

```{r ANOVA + Tukey urban filter}
#| warning: false

summary(aov(Total ~ factor(Zones),data = Traps_snake))

TukeyHSD(aov(Total ~ factor(Zones),data = Traps_snake))
```

# Source-sink dynamic model

```{r source-sink dynamic}
#| warning: false
# eval: false --> if we don't want to execute this block

source_sink_plot <- function(){
  
  # Visualization plot
  lizard_multiplot <- function(n_rep = 10, # Number of replicates
                               initial_pop = 300, # Initial number of individuals
                               n_generations = 41, # Number of generations the simulation is going to last
                               cat_pred = 0.10, # % of the total population that anthropogenic activities kill every year
                               snake_pred = 0.50, # % of the total population that snakes kill every year
                               snake_intro = 21, # Year that the snake has been introduced into the ecosystem
                               k = 300, # Carrying capacity
                               urban_index = 0, # Urbanization index
                               emigration_perc = 0 ){ # % of individuals of the carrying capacity that inmigrate
    
    # Create data frame to store number of individuals of each population during each simulation, together with the average number of individuals per generation
    all_simulations <- as.data.frame(matrix(NA, nrow = n_rep + 1, ncol = n_generations))
    
    # Create vector to store the generation number of the first extinct population
    extinction_gens <- rep(NA, n_rep)
    
    for(x in 1:n_rep){
      
      # Execute simulation x times
      lizard_projection_no_plot(initial_pop = initial_pop, 
                                n_generations = n_generations, 
                                cat_pred = cat_pred, 
                                snake_pred = snake_pred, 
                                snake_intro = snake_intro, 
                                k = k, 
                                urban_index = urban_index,
                                emigration_perc = emigration_perc)
      
      # Store number of individuals to the correspondant row
      all_simulations[x, ] <- n_ind_pop
      
      # Store the generation number where each population has become extinct (if it applies)
      extinction_gen <- which(n_ind_pop <= 0)[1]
      if (!is.na(extinction_gen)) {
        extinction_gens[x] <- extinction_gen
      }
    }
    
    # Average number of individuals per generation
    all_simulations[n_rep + 1, ] <- colMeans(all_simulations[1:n_rep, ])
    
    # Find the generation number of the first population that has become extinct
    first_extinction_gen <- min(extinction_gens, na.rm = TRUE)
    
    # Find the generation number of the last population that has become extinct
    avg_extinction_gen <- which(all_simulations[n_rep + 1, ] <= 0)[1]
    
    # Find the generation number where 50% of the populations are extinct
    half_extinction_gen <- NA
    for (gen in 1:n_generations) {
      if (sum(all_simulations[1:n_rep, gen] <= 0) >= n_rep / 2) {
        half_extinction_gen <- gen
        break
      }
    }
    
    assign("all_simulations", all_simulations, envir=globalenv())
    
    # Plot number of individuals per generation of all simulations
    plot(NULL, xlim = c(1, n_generations), ylim = c(0, (k + 50)), xlab = "Years", ylab = "Nº ind", xaxt = "n", cex.lab = 1.5, cex.axis = 1.5)
    axis(1, at = c(1, 11, 21, 31, 41, 51, 61), labels = c("-40", "-30", "-20", "-10", "0", "+10", "+20"), cex.axis = 1.5)
    
    for(x in 1:n_rep){
      lines(1:n_generations, all_simulations[x, ], type = "l", col = "grey", lwd = 1)
    }
    
    lines(1:n_generations, all_simulations[n_rep + 1, ], type = "l", col = "black", lwd = 3)
    
    # Draw red line when the snake is introduced into the ecosystem
    abline(v = snake_intro, col = "red3", lty = "solid")
    
    # Add vertical dashed line for all 3 types of extinction (first, 50%, and last)
    if (!is.na(first_extinction_gen)) {
      abline(v = first_extinction_gen, col = "grey60", lty = "dashed")
    }
    
    if (!is.na(half_extinction_gen)) {
      abline(v = half_extinction_gen, col = "grey40", lty = "dotdash")
    }
    
    if (!is.na(avg_extinction_gen)) {
      abline(v = avg_extinction_gen, col = "grey20", lty = "longdash")
    }
  }
  
  # Popuation simulation
  lizard_projection_no_plot <- function(initial_pop = 100, n_generations = 10, cat_pred = 0.10, snake_pred = 0.20, snake_intro = 10, k = 312, urban_index = 1, emigration_perc = 0.1) {
    
    # Initial population and array
    n_lizards <- initial_pop
    lizard_pop <- array(data = NA, dim = c(999999, 4, n_generations))
    dimnames(lizard_pop)[[2]] <- c("ID", "Age", "Sex", "Status")
    
    # First individuals
    for (i in 1:n_lizards) {
      lizard_pop[i, 1, 1] <- i # Lizard ID
      lizard_pop[i, 2, 1] <- pmin(rgeom(1, 0.2), 18) # Age (coming from a geometric distribution)
      lizard_pop[i, 3, 1] <- sample(c("Male", "Female"), 1) # Sex
      lizard_pop[i, 4, 1] <- 1 # Dead (0) or alive (1)
    }
    
    # Time simulation
    for (i in 2:n_generations) {
      
      # Keep tracking of number of rows filled
      last_ind <- sum(complete.cases(lizard_pop[, 1, i - 1]))
      
      # Select previous generation
      lizards_generation <- as.data.frame(lizard_pop[, , i - 1])
      
      # Select only alive individuals
      lizards_alive <- subset(lizards_generation, as.numeric(Status) == 1)
      
      # Anthropogenic factors mortality
      if (nrow(lizards_alive) > 0) {
        dead_ind_cat <- sample(1:nrow(lizards_alive), trunc(nrow(lizards_alive) * cat_pred), replace = FALSE)
        dead_ids_cat <- lizards_alive[dead_ind_cat, "ID"]
        for (id in dead_ids_cat) {
          idx <- which(lizard_pop[, 1, i - 1] == id)
          lizard_pop[idx, , i] <- lizard_pop[idx, , i - 1]
          lizard_pop[idx, 4, i] <- 0
        }
      }
      
      # Select individuals still alive after anthropogenic mortality
      lizards_alive_after_cats <- subset(as.data.frame(lizard_pop[, , i - 1]), as.numeric(Status) == 1 & is.na(lizard_pop[, 4, i]))
      
      # Snake mortality (starting from snake_intro)
      # Full urbanization counters snake effect
      if (i >= (snake_intro + 1)) {
        if (nrow(lizards_alive_after_cats) > 0) {
          dead_ind_snake <- sample(1:nrow(lizards_alive_after_cats), trunc(nrow(lizards_alive_after_cats) * (snake_pred - (urban_index * snake_pred))), replace = FALSE)
          dead_ids_snake <- lizards_alive_after_cats[dead_ind_snake, "ID"]
          for (id in dead_ids_snake) {
            idx <- which(lizard_pop[, 1, i - 1] == id)
            lizard_pop[idx, , i] <- lizard_pop[idx, , i - 1]
            lizard_pop[idx, 4, i] <- 0
          }
        }
      }
      
      # Select individuals still alive after snake mortality
      lizards_alive_after_snakes <- subset(as.data.frame(lizard_pop[, , i - 1]), as.numeric(Status) == 1 & is.na(lizard_pop[, 4, i]))
      
      # Denso dependency
      n_alive <- nrow(lizards_alive_after_snakes)
      
      if (n_alive > k) {
        surplus <- n_alive - k
        surplus_ind <- sample(1:nrow(lizards_alive_after_snakes), surplus, replace = FALSE)
        surplus_ids <- lizards_alive_after_snakes[surplus_ind, "ID"]
        for (id in surplus_ids) {
          idx <- which(lizard_pop[, 1, i - 1] == id)
          lizard_pop[idx, , i] <- lizard_pop[idx, , i - 1]
          lizard_pop[idx, 4, i] <- 0
        }
      }
      
      # The rest of individuals
      for (j in 1:last_ind) {
        
        # If there is an NA, jump to next loop
        if (is.na(lizard_pop[j, 4, i - 1])) {
          next
        }
        
        # If the individual is already dead in this generation (external death causes), jump to the next loop
        if (!is.na(lizard_pop[j, 4, i]) && as.numeric(lizard_pop[j, 4, i]) == 0) {
          next
        }
        
        # If lizard was dead in the previous generation, mark it as dead
        if (as.numeric(lizard_pop[j, 4, i - 1]) == 0) {
          lizard_pop[j, , i] <- lizard_pop[j, , i - 1]
          next
        }
        
        # If the lizard is alive, but it's older than 18 years old, it dies
        if (as.numeric(lizard_pop[j, 2, i - 1]) == 18) {
          lizard_pop[j, , i] <- lizard_pop[j, , i - 1]
          lizard_pop[j, 4, i] <- 0
          next
        }
        
        # Intrinsic survival probability
        surv_prob <- runif(1, 0.75, 0.83)
        surv_random <- runif(1, 0, 1)
        
        # If it fails, the lizard is marked as dead
        if (surv_random > surv_prob) {
          lizard_pop[j, , i] <- lizard_pop[j, , i - 1]
          lizard_pop[j, 4, i] <- 0
          
        } else {
          # If the lizard survives, it gets older and survives
          lizard_pop[j, 1, i] <- lizard_pop[j, 1, i - 1]
          lizard_pop[j, 2, i] <- as.numeric(lizard_pop[j, 2, i - 1]) + 1
          lizard_pop[j, 3, i] <- lizard_pop[j, 3, i - 1]
          lizard_pop[j, 4, i] <- 1
        }
        
        # Reproduction
        # If the individual is a female, older than 2 years old and alive
        if (lizard_pop[j, 3, i - 1] == "Female" && as.numeric(lizard_pop[j, 4, i - 1]) == 1 && as.numeric(lizard_pop[j, 2, i - 1]) >= 2) {
          
          if (as.numeric(lizard_pop[j, 2, i - 1]) > (18 + 1)) {
            # If it's an old female, it lays eggs two times (for the moment, only one egg laying per year)
            
            for (p in 1:2) {
              
              # Number of eggs laid
              n_eggs <- trunc(abs(rnorm(1, 2.30, 0.55)))
              
              if (n_eggs > 0) {
                
                for (m in 1:n_eggs) {
                  
                  # Hatching probability
                  hatch_prob <- runif(1, 0.60, 0.75)
                  hatch_random <- runif(1, 0, 1)
                  
                  # If the egg hatches
                  if (hatch_random < hatch_prob) {
                    
                    # Add a new row
                    last_ind <- last_ind + 1
                    lizard_pop[last_ind, 1, i] <- last_ind
                    lizard_pop[last_ind, 2, i] <- 0
                    lizard_pop[last_ind, 3, i] <- sample(c("Male", "Female"), 1)
                    lizard_pop[last_ind, 4, i] <- 1
                  }
                }
              }
            }
            
          } else {
            
            # If the female is young, only one laying
            n_eggs <- trunc(abs(rnorm(1, 2.30, 0.55)))
            
            if (n_eggs > 0) {
              
              for (m in 1:n_eggs) {
                
                hatch_prob <- runif(1, 0.60, 0.75)
                hatch_random <- runif(1, 0, 1)
                
                if (hatch_random < hatch_prob) {
                  
                  last_ind <- last_ind + 1
                  lizard_pop[last_ind, 1, i] <- last_ind
                  lizard_pop[last_ind, 2, i] <- 0
                  lizard_pop[last_ind, 3, i] <- sample(c("Male", "Female"), 1)
                  lizard_pop[last_ind, 4, i] <- 1
                }
              }
            }
          }
        }
      }
      
      # Emigration
      emigration <- trunc(k * emigration_perc)
      
      for (e in 1:emigration) {
        last_ind <- last_ind + 1
        lizard_pop[last_ind, 1, i] <- last_ind
        lizard_pop[last_ind, 2, i] <- pmin(rgeom(1, 0.2), 18) 
        lizard_pop[last_ind, 3, i] <- sample(c("Male", "Female"), 1)
        lizard_pop[last_ind, 4, i] <- 1
      }
      
      # Check for virtual extinction
      n_alive <- sum(lizard_pop[, 4, i] == 1, na.rm = TRUE)
      if (n_alive <= (k * 0.1)) { # if population is under 10% of carrying capacity, it becomes virtually extinct
        extinction_generation <- i
        for (j in 1:last_ind) {
          if (lizard_pop[j, 4, i] == 1) {
            lizard_pop[j, 4, i:n_generations] <- 0
          }
        }
        break
      }
      
      assign("lizard_pop", lizard_pop, envir = globalenv())
      
    }
    
    # Track nº of individuals per generation
    n_ind_pop <- as.data.frame(matrix(data = NA, ncol = n_generations, nrow = 1))
    for (i in 1:n_generations) {
      n_ind_pop[i] <- sum(lizard_pop[, 4, i] == 1, na.rm = TRUE)
    }
    
    colnames(n_ind_pop) <- paste0("Generation_", 1:n_generations)
    
    assign("n_ind_pop", n_ind_pop, envir=globalenv())
    
  }
  
  
  # Set layout for the final plot
  par(mfrow = c(2,6))
  
  
  # Simulation 1
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.05, 
                   snake_pred = 0.0, 
                   snake_intro = 70, 
                   k = 300, 
                   urban_index = 1, 
                   emigration_perc = 0) 
  
  # Simulation 2
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61,
                   cat_pred = 0.075, 
                   snake_pred = 0.0, 
                   snake_intro = 70, 
                   k = 300, 
                   urban_index = 1, 
                   emigration_perc = 0)
  # Simulation 3
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.1, 
                   snake_pred = 0.0, 
                   snake_intro = 70, 
                   k = 300, 
                   urban_index = 1, 
                   emigration_perc = 0) 
  # Simulation 4
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61,
                   cat_pred = 0.125, 
                   snake_pred = 0.0, 
                   snake_intro = 70, 
                   k = 300, 
                   urban_index = 1, 
                   emigration_perc = 0)
  # Simulation 5
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.15, 
                   snake_pred = 0.0, 
                   snake_intro = 70, 
                   k = 300, 
                   urban_index = 1, 
                   emigration_perc = 0) 
  # Simulation 6
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.25, 
                   snake_pred = 0.0, 
                   snake_intro = 70, 
                   k = 300, 
                   urban_index = 1, 
                   emigration_perc = 0) 
  # Simulation 7
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.10, 
                   snake_pred = 0.25, 
                   snake_intro = 21, 
                   k = 300, 
                   urban_index = 0, 
                   emigration_perc = 0) 
  # Simulation 8
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.10, 
                   snake_pred = 0.25, 
                   snake_intro = 21, 
                   k = 300, 
                   urban_index = 0.25, 
                   emigration_perc = 0) 
  # Simulation 9
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.10, 
                   snake_pred = 0.25, 
                   snake_intro = 21, 
                   k = 300, 
                   urban_index = 0.5, 
                   emigration_perc = 0) 
  # Simulation 10
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.10, 
                   snake_pred = 0.25, 
                   snake_intro = 21, 
                   k = 300, 
                   urban_index = 0.8, 
                   emigration_perc = 0) 
  # Simulation 11
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.10, 
                   snake_pred = 0.25, 
                   snake_intro = 21, 
                   k = 300, 
                   urban_index = 0.9, 
                   emigration_perc = 0) 
  
  # Simulation 12
  lizard_multiplot(n_rep = 10, 
                   initial_pop = 300, 
                   n_generations = 61, 
                   cat_pred = 0.10, 
                   snake_pred = 0.25, 
                   snake_intro = 21, 
                   k = 300, 
                   urban_index = 1, 
                   emigration_perc = 0) 
  
}

source_sink_plot()
```
